---
title: "Where irrigation exists is globally contested"
subtitle: "5. Analyses supporting the reply to the reviewers"
author: "Arnald Puy"
header-includes:
  - \usepackage[font=footnotesize]{caption}
  - \usepackage{dirtytalk}
  - \usepackage{booktabs}
  - \usepackage{tabulary}
  - \usepackage{enumitem}
  - \usepackage{lmodern}
  - \usepackage{amsmath}
  - \usepackage{mathtools}
  - \usepackage[T1]{fontenc}
  - \usepackage{tikz}
output:
  pdf_document:
    fig_caption: yes
    number_sections: yes
    toc: yes
    toc_depth: 2
    keep_tex: true
  word_document:
    toc: no
    toc_depth: '2'
  html_document:
    keep_md: true
link-citations: yes
fontsize: 11pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, dev = "pdf", cache = TRUE)
```

\newpage

# Preliminary

```{r preliminary, warning=FALSE, message=FALSE, results = "hide"}

# ANALYSES SUPPORTING THE REPLY TO THE REVIEWERS ##############################
################################################################################

# This script reproduces every quantitative figure cited in our reply to the
# reviewers. All analyses use the harmonized ensemble at 0.2 deg restricted to
# cropland cells (the same universe as the main analysis).

# Load libraries ---------------------------------------------------------------

sensobol::load_packages(c("data.table", "here", "terra", "magrittr"))

# Functions reused from the main analysis --------------------------------------

source(here("functions", "filter_long_dt_by_crop_mask_aggregated.R"))
source(here("functions", "compute_tau_max_disagreement_fun.R"))

# Harmonized ensemble at 0.2 deg, restricted to cropland -----------------------

dt <- fread(here("datasets", "irrigated_areas_regridded",
                 "irrigated_areas_regridded_02.csv"))
dt[, resolution := "0.2deg"]
crop <- rast(here("irrigated_area_datasets", "GirrEO", "asap_mask_crop_v02.tif"))
dt <- filter_long_dt_by_crop_mask_aggregated(dt = dt, crop_native = crop,
                                             rule = "any")

# Cells x maps -----------------------------------------------------------------

wide <- dcast(dt, lon + lat + country + continent ~ dataset, value.var = "mha",
              fill = 0)

maps <- c("gaez_v4", "giam", "gmia", "gripc", "luh2", "meier",
          "mirca2000", "mirca_os", "nagaraj", "spam")
stopifnot(all(maps %in% names(wide)))

# Nominal reference year of each map (Table on temporal characteristics) -------

nominal <- c(giam = 2000, mirca2000 = 2000, gmia = 2005, gripc = 2005,
             nagaraj = 2005, meier = 2005, spam = 2010, mirca_os = 2015,
             gaez_v4 = 2015, luh2 = 2015)

# Mapping families -------------------------------------------------------------

census <- c("gmia", "mirca2000", "mirca_os", "spam", "gaez_v4")  # FAO-statistics
rs      <- c("giam", "gripc")                                    # remote sensing

# Detectability threshold grid (1 ha to 1e5 ha) and 1%-of-cell tolerance -------

taus_ha     <- c(0, 10^(seq(0, 5, by = 0.1)))
taus_mha    <- taus_ha * 1e-6
zero_tol_02 <- 500  # 1% of the 0.2 deg cell, in ha

# Existential disagreement: of the cells where at least one map records
# irrigation (> 0), the fraction in which at least one map records exactly zero.

exist_union <- function(w, cols) {
  A <- as.matrix(w[, ..cols])
  n_present <- rowSums(A > 0, na.rm = TRUE)
  disagree  <- n_present > 0 & n_present < length(cols)
  Amin <- do.call(pmin, c(as.data.frame(A), list(na.rm = TRUE)))
  sum(disagree & Amin == 0) / sum(n_present > 0)
}

# Extreme disagreement: fraction of cells whose tau_max exceeds ten times the
# detectability threshold (one map reports > ~10% of the cell, another ~0).

extreme_frac <- function(w, cols, zero_tol = zero_tol_02) {
  tm <- compute_tau_max_existential_fun(w, cols, taus_mha, taus_ha,
                                        zero_tol_ha = zero_tol)
  mean(!is.na(tm$tau_max_ha) & tm$tau_max_ha > 10 * zero_tol)
}
```

\newpage

# Reviewer 1

## R1.1: the detectability threshold as a fraction of the grid cell

The same absolute $\tau$ (in hectares) is a different fraction of the grid cell
at each resolution. We report the equator cell areas and translate key $\tau$
values into cell fractions.

```{r r11}

# Equator cell area (ha) per resolution ----------------------------------------

R_earth <- 6371
deg <- c("0.2" = 0.2, "0.4" = 0.4, "1" = 1)
cell_ha <- sapply(deg, function(d) (d * pi / 180 * R_earth)^2 * 100)
round(cell_ha)

# Key tau values as a percentage of the cell ----------------------------------

tau_key <- c(100, 500, 1000, 2500, 5000, 10000, 25000)
tau_tab <- data.table(
  `tau (ha)`    = tau_key,
  `0.2 deg (%)` = round(100 * tau_key / cell_ha["0.2"], 2),
  `0.4 deg (%)` = round(100 * tau_key / cell_ha["0.4"], 2),
  `1 deg (%)`   = round(100 * tau_key / cell_ha["1"],   2))

knitr::kable(tau_tab, caption = "Detectability threshold tau as a percentage of the grid-cell area at the equator. The analysis centres on tau = 1% of the cell (about 500, 2000 and 12400 ha); extreme disagreement denotes tau_max above ten times this value.")
```

## R1.5: temporal heterogeneity does not explain the disagreement

We test whether the disagreement reflects the different moments the maps
represent. Existence of irrigation is the metric throughout (see Preliminary).

```{r r15_sameyear}

# Maps that share the 2005 reference year versus the full ensemble --------------

sameyear_2005 <- c("gmia", "gripc", "nagaraj", "meier")

data.table(
  set  = c("Same year (2005)", "Full ensemble"),
  maps = c(length(sameyear_2005), length(maps)),
  existential_pct = round(100 * c(exist_union(wide, sameyear_2005),
                                  exist_union(wide, maps))))
```

```{r r15_sizecontrol}

# Recompute over every four-map combination to hold ensemble size fixed --------

combs4  <- combn(maps, 4, simplify = FALSE)
ev4     <- sapply(combs4, function(cc) exist_union(wide, cc))
span4   <- sapply(combs4, function(cc) diff(range(nominal[cc])))
same_ev <- exist_union(wide, sameyear_2005)

data.table(
  quantity = c("Minimum", "Median", "Maximum",
               "Same-year (2005) combination", "Most dispersed (1999-2015)"),
  existential_pct = round(100 * c(min(ev4), median(ev4), max(ev4),
                                  same_ev, mean(ev4[span4 == max(span4)]))),
  percentile = c(NA, NA, NA, round(100 * mean(ev4 <= same_ev)), NA))
```

```{r r15_directional}

# Direction of disagreement: under real expansion, a recent (2015) map should
# not record zero where an older (2000-2005) map detects irrigation.

old_t    <- c("giam", "mirca2000", "gmia", "gripc")   # 2000-2005
recent_t <- c("gaez_v4", "luh2", "mirca_os")          # 2015

pooled_blind <- function(new) {
  A_old   <- as.matrix(wide[, ..old_t])
  old_any <- rowSums(A_old > 0, na.rm = TRUE) > 0
  sum(old_any & wide[[new]] == 0, na.rm = TRUE) / sum(old_any)
}

data.table(
  recent_map = c("GAEZ+2015", "LUH2", "MIRCA-OS"),
  sees_zero_where_older_detects_pct = round(100 * sapply(recent_t, pooled_blind)))
```

\newpage

# Reviewer 2

## R2.1: census products agree because they share source data, not reality

The five FAO-statistics products share definitions and much of their source
data. We compare their existential disagreement against all five-map subsets.

```{r r21}

census_ev <- exist_union(wide, census)

combs5 <- combn(maps, 5, simplify = FALSE)
ev5    <- sapply(combs5, function(cc) exist_union(wide, cc))

data.table(
  quantity = c("Census-5 (GMIA, MIRCA2000, MIRCA-OS, SPAM, GAEZ+2015)",
               "All 252 five-map subsets: minimum",
               "All 252 five-map subsets: median",
               "All 252 five-map subsets: maximum"),
  existential_pct = round(100 * c(census_ev, min(ev5), median(ev5), max(ev5))),
  note = c(sprintf("percentile %.0f%%", 100 * mean(ev5 <= census_ev)),
           "", "", ""))
```

## R2.2: remote-sensing and census products are not a subset of one another

If remote sensing simply mapped a superset of census irrigation, it would never
record zero where a census product detects irrigation. We test both directions
and the disagreement within each family.

```{r r22}

any_pos <- function(cols) rowSums(as.matrix(wide[, ..cols]) > 0, na.rm = TRUE) > 0
cP <- any_pos(census)
rP <- any_pos(rs)

data.table(
  quantity = c("RS records zero where census detects irrigation (% of census-irrigated cells)",
               "Census records zero where RS detects irrigation (% of RS-irrigated cells)",
               "Within RS family: GIAM vs GRIPC existential disagreement (%)",
               "Within census family: five FAO products existential disagreement (%)",
               "Mean global total, census maps (Mha)",
               "Mean global total, RS maps (Mha)"),
  value = c(round(100 * sum(cP & !rP) / sum(cP)),
            round(100 * sum(rP & !cP) / sum(rP)),
            round(100 * exist_union(wide, rs)),
            round(100 * exist_union(wide, census)),
            round(mean(sapply(census, function(m) sum(wide[[m]])))),
            round(mean(sapply(rs,     function(m) sum(wide[[m]]))))))

# Per-map global totals underlying the census/RS means (which maps each group
# uses follows the reviewer's own classification in his comment) ---------------

data.table(
  group = c(rep("census", length(census)), rep("remote sensing", length(rs))),
  map = c(census, rs),
  total_mha = round(sapply(c(census, rs), function(m) sum(wide[[m]])), 1))
```

## R2.3: census products are compared on a common physical extent

The reviewer notes that the census products map different quantities (area
equipped, harvested area, monthly growing area). We do not compare these raw
quantities: in `code_original_datasets` we reduce every product to its actual
physical irrigated extent, so these magnitude differences do not enter the
comparison. The variable used for each product is recorded below.

```{r r23}

provenance <- data.table(
  product = c("GMIA", "SPAM2010", "MIRCA2000", "MIRCA-OS", "GAEZ+2015", "GIAM"),
  variable_used = c("Actually irrigated area (AEI x AAI fraction), not AEI",
                    "Physical irrigated area (SPAM 'A'), capped at cell area",
                    "Maximum monthly irrigated area (physical extent), not summed",
                    "Maximum monthly growing area (physical extent), not summed",
                    "Harvested irrigated area, capped at physical cell area",
                    "Net actually irrigated area (TAAI), not annualized (AIA)"),
  source = c("code_original_datasets.Rmd L254", "L439, L452", "L341", "L643",
             "L382, L399", "L205"))

knitr::kable(provenance, caption = "Variable extracted for each product (code_original_datasets.Rmd). Every product is reduced to its actual physical irrigated extent, so differences between area equipped, harvested area and monthly growing area do not enter the existence comparison.")
```

\newpage

# Reviewer 3

## R3.2: where the maps agree on the existence of irrigation

The reviewer asks where all datasets agree on irrigation presence. We identify
cells where all ten maps detect irrigation above the detectability threshold
(1% of the cell, 500 ha at 0.2 deg) -- the "agreement on presence" of the
tau_max framework -- and characterise them, including their overlap with rice
production (EarthStat).

```{r r32}

Apres <- as.matrix(wide[, ..maps])
wide[, agree_presence := rowSums(Apres > zero_tol_02 * 1e-6) == 10]

# Overlap with rice production (EarthStat, regridded to 0.2 deg) ---------------

rice <- rast(here("datasets", "earthstat_regridded", "res_02",
                  "rice_Production_ll_02.tif"))
wide[, rice_pos := { v <- terra::extract(rice, cbind(lon, lat))[, 1]
                     !is.na(v) & v > 0 }]

ag <- wide[agree_presence == TRUE]

data.table(
  metric = c("Agreement cells (all 10 above 1% of cell)",
             "In Asia (%)",
             "Latitude belt (10th-90th percentile)",
             "Coincide with rice production (%)",
             "All cropland cells coinciding with rice (%)"),
  value = c(as.character(nrow(ag)),
            as.character(round(100 * mean(ag$continent == "Asia"))),
            sprintf("%.0f-%.0f N", quantile(ag$lat, .10), quantile(ag$lat, .90)),
            as.character(round(100 * mean(ag$rice_pos))),
            as.character(round(100 * mean(wide$rice_pos)))))
```

```{r r32_countries}

knitr::kable(ag[, .N, country][order(-N)][1:5],
             caption = "Countries with the most cells where all ten datasets agree irrigation is present (0.2 deg).")
```

## R3.14: disagreement does not track irrigation practice

The reviewer asks whether irrigation technique (e.g. drip in Spain) drives the
disagreement. We report existential and extreme disagreement per country.

```{r r314}

# Global reference -------------------------------------------------------------

glob <- data.table(country = "GLOBAL", cells = nrow(wide),
                   existential_pct = round(100 * exist_union(wide, maps)),
                   extreme_pct = round(100 * extreme_frac(wide, maps)))

# By country -------------------------------------------------------------------

ctys <- c("India", "Portugal", "Italy", "Pakistan", "China", "Spain",
          "Egypt", "France", "Mexico", "Iran", "United States")

res <- rbindlist(lapply(ctys, function(cn) {
  s <- wide[country == cn]
  if (nrow(s) == 0) return(NULL)
  data.table(country = cn, cells = nrow(s),
             existential_pct = round(100 * exist_union(s, maps)),
             extreme_pct = round(100 * extreme_frac(s, maps)))
}))

res <- rbind(glob, res[order(-extreme_pct)])

knitr::kable(res, caption = "Existential and extreme disagreement by country (0.2 deg). Existential: of the cells where at least one map detects irrigation, the percentage in which at least one map detects none. Extreme: percentage of cropland cells whose tau_max exceeds ten times the detectability threshold. Spain (the highest-drip country) is not a maximum.")
```

\newpage

# Additional analyses: the non-identifiability claim

These analyses support the manuscript section "Irrigated areas are
non-identifiable". The rationale and interpretation are given there and in the
SM (Secs. 1.7.3-1.7.5); here we only reproduce the numbers.

## Underdetermination: how many truths are consistent with the data

Determined-present: every map exceeds the strictest admissible threshold
($\min_d A_d > T$). Determined-absent: every map reports exactly zero (none
here by construction, since the universe only contains cells with at least one
positive report). All other cells are underdetermined: both classifications
are consistent with the ensemble.

```{r underdetermination}

# UNDERDETERMINATION COUNT #####################################################

A <- as.matrix(wide[, ..maps])
Amin <- do.call(pmin, as.data.frame(A))
Amax <- do.call(pmax, as.data.frame(A))

underdet_fun <- function(T_ha) {
  T_mha    <- T_ha * 1e-6
  det_pres <- Amin > T_mha
  und      <- !(det_pres | Amax == 0)
  data.table(T_ha             = T_ha,
             det_present      = sum(det_pres),
             underdetermined  = sum(und),
             underdet_pct     = round(100 * mean(und), 1),
             existential_pct  = round(100 * sum(und & Amin == 0) / sum(und), 1),
             log10_truths     = round(sum(und) * log10(2)))
}

knitr::kable(rbindlist(lapply(c(100, 500, 1000), underdet_fun)),
             caption = "Underdetermination of irrigation presence (0.2 deg,
             cropland). det\\_present: cells where every map exceeds the
             strictest admissible threshold T. existential\\_pct: share of
             underdetermined cells where at least one map reports exactly
             zero. log10\\_truths: size (in powers of ten) of the set of
             complete global presence fields consistent with all ten maps.")
```

Consistency checks: determined-present at $T=500$ ha (9,248 cells) equals the
R3.2 agreement-cell count; the existential set (117,294 cells, 90.5\%) equals
the manuscript's 90\% headline.

## Latent-class analysis: testing "identifiable truth plus noise"

$K$-class latent class models (conditional independence within class) fitted
by expectation-maximisation on the $2^{10}$ response patterns of the binary
presence matrix, at $\tau=0$ and $\tau=500$ ha, with 50 random starts per fit
(Hui and Walter 1980).

```{r lca_functions}

# K-CLASS LATENT CLASS MODEL VIA EM ON AGGREGATED RESPONSE PATTERNS ############

lca_em <- function(pat, cnt, K, n_start = 50, tol = 1e-8, maxit = 3000,
                   seed = 123) {
  set.seed(seed)
  D <- ncol(pat); n <- sum(cnt)
  best <- NULL; logliks <- numeric(n_start)
  for (s in seq_len(n_start)) {
    x    <- rexp(K); pi_k <- x / sum(x)
    rho  <- matrix(runif(K * D, .05, .95), K, D)
    ll_old <- -Inf
    for (it in seq_len(maxit)) {
      logp <- sapply(seq_len(K), function(k)
        pat %*% log(rho[k, ]) + (1 - pat) %*% log(1 - rho[k, ]) + log(pi_k[k]))
      m    <- apply(logp, 1, max)
      lse  <- m + log(rowSums(exp(logp - m)))
      ll   <- sum(cnt * lse)
      post <- exp(logp - lse)
      w    <- post * cnt
      Nk   <- colSums(w)
      pi_k <- Nk / n
      rho  <- pmin(pmax(crossprod(w, pat) / Nk, 1e-6), 1 - 1e-6)
      if (abs(ll - ll_old) < tol) break
      ll_old <- ll
    }
    logliks[s] <- ll
    if (is.null(best) || ll > best$ll)
      best <- list(ll = ll, pi = pi_k, rho = rho, post = post)
  }
  npar <- (K - 1) + K * D
  c(best, list(npar = npar, bic = -2 * best$ll + npar * log(n),
               logliks = logliks))
}

# Fit K = 1-4, report fit, error rates, residual dependence, 3-class profile --

lca_report_fun <- function(Y, label) {
  D    <- ncol(Y); n <- nrow(Y)
  code <- as.integer(Y %*% 2^(0:(D - 1)))
  tab  <- data.table(code = code)[, .N, by = code][order(code)]
  pat  <- t(sapply(tab$code, function(z) as.integer(intToBits(z)[1:D])))
  cnt  <- tab$N
  fits <- lapply(1:4, function(K) lca_em(pat, cnt, K))

  comp <- data.table(K = 1:4,
                     loglik = sapply(fits, function(f) round(f$ll, 1)),
                     npar   = sapply(fits, function(f) f$npar),
                     BIC    = sapply(fits, function(f) round(f$bic, 1)))

  f2  <- fits[[2]]
  irr <- which.max(rowMeans(f2$rho)); no <- setdiff(1:2, irr)
  err <- data.table(map = maps,
                    FNR = round(1 - f2$rho[irr, ], 3),
                    FPR = round(f2$rho[no, ], 3))[order(-FNR)]

  logp <- sapply(1:2, function(k)
    pat %*% log(f2$rho[k, ]) + (1 - pat) %*% log(1 - f2$rho[k, ]) +
      log(f2$pi[k]))
  m  <- apply(logp, 1, max)
  e  <- n * exp(m + log(rowSums(exp(logp - m))))
  G2 <- 2 * sum(cnt * log(cnt / e))
  df <- nrow(pat) - 1 - f2$npar

  P_obs <- crossprod(Y) / n
  P_exp <- f2$pi[1] * tcrossprod(f2$rho[1, ]) +
           f2$pi[2] * tcrossprod(f2$rho[2, ])
  R  <- (P_obs - P_exp) * 100
  ut <- which(upper.tri(R), arr.ind = TRUE)
  res <- data.table(m1 = maps[ut[, 1]], m2 = maps[ut[, 2]],
                    resid = round(R[ut], 2))[order(-abs(resid))][1:5]

  f3  <- fits[[3]]
  ord <- order(rowMeans(f3$rho))
  pr  <- data.table(class = sprintf("pi = %.2f", f3$pi[ord]),
                    round(as.data.table(f3$rho[ord, , drop = FALSE]), 2))
  setnames(pr, c("class", maps))
  post3 <- f3$post[match(code, tab$code), ]
  mid_share <- 100 * mean(apply(post3, 1, which.max) == ord[2])
  ll_r  <- round(f2$logliks, 1)

  cat("\n####", label, "----", n, "cells,", nrow(pat), "patterns\n")
  print(comp)
  cat(sprintf("\n2-class fit: G2 = %.0f on df = %d (G2/df = %.1f)\n",
              G2, df, G2 / df))
  cat(sprintf("2-class multistart: %d distinct loglik modes in 50 starts\n",
              length(unique(ll_r))))
  cat("\nImplied per-map error rates (2-class):\n"); print(err)
  cat("\nLargest residual pairwise associations (obs - exp P(both = 1), x100):\n")
  print(res)
  cat(sprintf("\n3-class model: BIC improvement over 2-class = %.0f;",
              fits[[2]]$bic - f3$bic))
  cat(sprintf(" middle-class posterior share = %.1f%%\n", mid_share))
  cat("3-class item-probability profiles (rows ordered low to high):\n")
  print(pr)
  invisible(fits)
}
```

```{r lca_run}

# RUN AT BOTH DETECTION CONVENTIONS ############################################

fits_tau0   <- lca_report_fun(1L * (A > 0),    "tau = 0 (any positive area)")
fits_tau500 <- lca_report_fun(1L * (A > 5e-4), "tau = 500 ha (1% of cell)")
```

## The best-products test over the United States

LANID (Xie and Lark 2021; 30 m, Landsat, annual), MIrAD-US v4 (Pervez and
Brown 2010; 250 m, MODIS anchored to USDA county statistics) and AIM-HPA
(Deines et al. 2019; 30 m, Landsat, High Plains Aquifer) compared in 2012 and
2017, the two years all three cover. Sources: LANID doi:10.5281/zenodo.5548555;
MIrAD-US doi:10.5066/P9NA3EO8; AIM-HPA
doi:10.4211/hs.a371fd69d41b4232806d81e17fe4efcb. The aggregation to the
0.2$^\circ$ grid (warp-sum of binary irrigated pixels $\times$ pixel area:
0.09 ha at 30 m, 6.25 ha at 250 m) is documented in the non-evaluated chunk
below; the aggregated table ships with the repository
(\texttt{us\_regional\_products/}). CONUS totals as a sanity check: LANID
23.1/24.2 Mha, MIrAD 22.5/23.8 Mha (2012/2017).

```{r us_test_aggregation, eval=FALSE}

# AGGREGATION OF THE US REGIONAL PRODUCTS (documentation; not run) #############

template <- rast(xmin = -126, xmax = -66, ymin = 24, ymax = 50,
                 resolution = 0.2, crs = "EPSG:4326")

agg_to_ha <- function(f, irr_val, px_ha) {
  ir <- (rast(f) == irr_val) * 1.0                # binary irrigated
  project(ir, template, method = "sum") * px_ha   # pixel count x pixel area
}

# lanid2012/2017.tif: 30 m, CONUS Albers, 0/1 (Zenodo record 5548555)
# mirad250_12/17v4.tif: 250 m, Lambert azimuthal EA, 0/1 (ScienceBase)
# {2012,2017}_AIM-HPA_rse_finalMapBinary.tif: 30 m; 0 = NoData, 1 = irrigated,
#   2 = not irrigated (HydroShare); validity = fraction of cell with value > 0
```

```{r us_test_metrics}

# EXISTENTIAL DISAGREEMENT: REGIONAL PRODUCTS VS THE GLOBAL ENSEMBLE ###########

# The CSV holds the product values extracted at the paper's 0.2 deg cell
# centres for every US cropland cell, so it joins back on lon/lat exactly
# (rounded to one decimal to be robust to floating-point noise).

reg <- fread(here("us_regional_products", "us_regional_products_02.csv"))
reg[, `:=`(lon = round(lon, 1), lat = round(lat, 1))]

us <- copy(wide[country == "United States"])
us[, `:=`(lon = round(lon, 1), lat = round(lat, 1))]
us <- merge(us, reg, by = c("lon", "lat"))
stopifnot(nrow(us) == nrow(reg))

# Existential: of union cells (>0 somewhere), % with at least one exact zero.
# Strong: % of union cells where one product is 0 and another exceeds 1% of
# the cell. Regional columns are in ha; the global maps are in Mha.

exist_pct <- function(d, cols) {
  A  <- as.matrix(d[, ..cols])
  np <- rowSums(A > 0); un <- np > 0
  round(100 * sum(un & np < length(cols)) / sum(un), 1)
}
strong_pct <- function(d, cols, big) {
  A  <- as.matrix(d[, ..cols])
  un <- rowSums(A > 0) > 0
  round(100 * sum(un & apply(A, 1, min) == 0 &
                    apply(A, 1, max) > big) / sum(un), 1)
}
dir_pct <- function(d, a, b) {           # % of union cells with a = 0, b > 0
  A  <- as.matrix(d[, c(a, b), with = FALSE])
  un <- rowSums(A > 0) > 0
  round(100 * sum(un & A[, 1] == 0 & A[, 2] > 0) / sum(un), 1)
}

hpa <- us[aim_valid > 0.5]

row_fun <- function(d, universe, comparison, cols, big)
  data.table(universe, comparison,
             existential = exist_pct(d, cols),
             strong      = strong_pct(d, cols, big))

out <- rbind(
  row_fun(us,  "CONUS",       "LANID x MIrAD 2012", c("lanid12", "mirad12"), 500),
  row_fun(us,  "CONUS",       "LANID x MIrAD 2017", c("lanid17", "mirad17"), 500),
  row_fun(us,  "CONUS",       "Ten global maps",    maps,                    5e-4),
  row_fun(hpa, "High Plains", "LANID x MIrAD 2012", c("lanid12", "mirad12"), 500),
  row_fun(hpa, "High Plains", "LANID x MIrAD 2017", c("lanid17", "mirad17"), 500),
  row_fun(hpa, "High Plains", "LANID x MIrAD x AIM 2012",
          c("lanid12", "mirad12", "aim12"), 500),
  row_fun(hpa, "High Plains", "LANID x MIrAD x AIM 2017",
          c("lanid17", "mirad17", "aim17"), 500),
  row_fun(hpa, "High Plains", "Ten global maps",    maps,                    5e-4))

knitr::kable(out, caption = "Existential disagreement (of the cells where at
             least one product records irrigation, the percentage where at
             least one records exactly zero) and strong contradictions (one
             product records zero, another exceeds 1\\% of the cell) among the
             best US regional irrigation products, benchmarked against the ten
             global maps over the same cells. CONUS: 17,087 US cropland cells;
             High Plains: 1,549 cells within the AIM-HPA footprint.")

# Directional asymmetry (mirrors the global census-vs-remote-sensing axis) -----

data.table(
  year = c(2012, 2017),
  lanid0_mirad_pos = c(dir_pct(us, "lanid12", "mirad12"),
                       dir_pct(us, "lanid17", "mirad17")),
  mirad0_lanid_pos = c(dir_pct(us, "mirad12", "lanid12"),
                       dir_pct(us, "mirad17", "lanid17")))
```

```{r us_test_structure}

# GEOGRAPHIC AND MAGNITUDE STRUCTURE OF THE LANID-MIRAD DISAGREEMENT ###########

# Directional cells by longitude (100W = the classic humid/arid divide) and
# the median irrigated area at stake in each direction.

struct <- rbindlist(lapply(c("12", "17"), function(yr) {
  L <- us[[paste0("lanid", yr)]]; M <- us[[paste0("mirad", yr)]]
  un <- L > 0 | M > 0
  lm <- un & L == 0 & M > 0; ml <- un & M == 0 & L > 0; both <- L > 0 & M > 0
  data.table(year = paste0("20", yr),
             direction = c("MIrAD > 0, LANID = 0", "LANID > 0, MIrAD = 0",
                           "Both > 0"),
             cells = c(sum(lm), sum(ml), sum(both)),
             east_100W_pct = round(100 * c(mean(us$lon[lm] > -100),
                                           mean(us$lon[ml] > -100),
                                           mean(us$lon[both] > -100))),
             median_ha = round(c(median(M[lm]), median(L[ml]),
                                 median(pmin(L, M)[both]))))
}))
struct
cat("Baseline: union cells east of 100W:",
    round(100 * mean(us$lon[us$lanid12 > 0 | us$mirad12 > 0] > -100)), "%\n")

# Pairwise existential disagreement within the High Plains -------------------

pair_exist <- function(a, b) {
  un <- a > 0 | b > 0
  round(100 * mean((a == 0 | b == 0)[un]), 1)
}
for (yr in c("12", "17"))
  cat(sprintf(
    "HPA 20%s pairwise existential: LANIDxMIrAD %.1f | LANIDxAIM %.1f | MIrADxAIM %.1f\n",
    yr,
    pair_exist(hpa[[paste0("lanid", yr)]], hpa[[paste0("mirad", yr)]]),
    pair_exist(hpa[[paste0("lanid", yr)]], hpa[[paste0("aim",   yr)]]),
    pair_exist(hpa[[paste0("mirad", yr)]], hpa[[paste0("aim",   yr)]])))
```

The interpretation of these results is given in the manuscript (section
"Irrigated areas are non-identifiable") and the SM (Sec. 1.7.5).

\newpage

# Session information

```{r session_information}

sessionInfo()
```
