---
title: "Where irrigation exists is globally contested"
subtitle: "5. Analyses supporting the reply to the reviewers"
author: "Arnald Puy"
header-includes:
  - \usepackage[font=footnotesize]{caption}
  - \usepackage{dirtytalk}
  - \usepackage{booktabs}
  - \usepackage{tabulary}
  - \usepackage{enumitem}
  - \usepackage{lmodern}
  - \usepackage{amsmath}
  - \usepackage{mathtools}
  - \usepackage[T1]{fontenc}
  - \usepackage{tikz}
output:
  pdf_document:
    fig_caption: yes
    number_sections: yes
    toc: yes
    toc_depth: 2
    keep_tex: true
  word_document:
    toc: no
    toc_depth: '2'
  html_document:
    keep_md: true
link-citations: yes
fontsize: 11pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, dev = "pdf", cache = TRUE)
```

\newpage

# Preliminary

```{r preliminary, warning=FALSE, message=FALSE, results = "hide"}

# ANALYSES SUPPORTING THE REPLY TO THE REVIEWERS ##############################
################################################################################

# This script reproduces every quantitative figure cited in our reply to the
# reviewers. All analyses use the harmonized ensemble at 0.2 deg restricted to
# cropland cells (the same universe as the main analysis).

# Load libraries ---------------------------------------------------------------

sensobol::load_packages(c("data.table", "here", "terra", "magrittr"))

# Functions reused from the main analysis --------------------------------------

source(here("functions", "filter_long_dt_by_crop_mask_aggregated.R"))
source(here("functions", "compute_tau_max_disagreement_fun.R"))

# Harmonized ensemble at 0.2 deg, restricted to cropland -----------------------

dt <- fread(here("datasets", "irrigated_areas_regridded",
                 "irrigated_areas_regridded_02.csv"))
dt[, resolution := "0.2deg"]
crop <- rast(here("irrigated_area_datasets", "GirrEO", "asap_mask_crop_v02.tif"))
dt <- filter_long_dt_by_crop_mask_aggregated(dt = dt, crop_native = crop,
                                             rule = "any")

# Cells x maps -----------------------------------------------------------------

wide <- dcast(dt, lon + lat + country ~ dataset, value.var = "mha", fill = 0)

maps <- c("gaez_v4", "giam", "gmia", "gripc", "luh2", "meier",
          "mirca2000", "mirca_os", "nagaraj", "spam")
stopifnot(all(maps %in% names(wide)))

# Nominal reference year of each map (Table on temporal characteristics) -------

nominal <- c(giam = 2000, mirca2000 = 2000, gmia = 2005, gripc = 2005,
             nagaraj = 2005, meier = 2005, spam = 2010, mirca_os = 2015,
             gaez_v4 = 2015, luh2 = 2015)

# Mapping families -------------------------------------------------------------

census <- c("gmia", "mirca2000", "mirca_os", "spam", "gaez_v4")  # FAO-statistics
rs      <- c("giam", "gripc")                                    # remote sensing

# Detectability threshold grid (1 ha to 1e5 ha) and 1%-of-cell tolerance -------

taus_ha     <- c(0, 10^(seq(0, 5, by = 0.1)))
taus_mha    <- taus_ha * 1e-6
zero_tol_02 <- 500  # 1% of the 0.2 deg cell, in ha

# Existential disagreement: of the cells where at least one map records
# irrigation (> 0), the fraction in which at least one map records exactly zero.

exist_union <- function(w, cols) {
  A <- as.matrix(w[, ..cols])
  n_present <- rowSums(A > 0, na.rm = TRUE)
  disagree  <- n_present > 0 & n_present < length(cols)
  Amin <- do.call(pmin, c(as.data.frame(A), list(na.rm = TRUE)))
  sum(disagree & Amin == 0) / sum(n_present > 0)
}

# Extreme disagreement: fraction of cells whose tau_max exceeds ten times the
# detectability threshold (one map reports > ~10% of the cell, another ~0).

extreme_frac <- function(w, cols, zero_tol = zero_tol_02) {
  tm <- compute_tau_max_existential_fun(w, cols, taus_mha, taus_ha,
                                        zero_tol_ha = zero_tol)
  mean(!is.na(tm$tau_max_ha) & tm$tau_max_ha > 10 * zero_tol)
}
```

\newpage

# Reviewer 1

## R1.1: the detectability threshold as a fraction of the grid cell

The same absolute $\tau$ (in hectares) is a different fraction of the grid cell
at each resolution. We report the equator cell areas and translate key $\tau$
values into cell fractions.

```{r r11}

# Equator cell area (ha) per resolution ----------------------------------------

R_earth <- 6371
deg <- c("0.2" = 0.2, "0.4" = 0.4, "1" = 1)
cell_ha <- sapply(deg, function(d) (d * pi / 180 * R_earth)^2 * 100)
round(cell_ha)

# Key tau values as a percentage of the cell ----------------------------------

tau_key <- c(100, 500, 1000, 2500, 5000, 10000, 25000)
tau_tab <- data.table(
  `tau (ha)`    = tau_key,
  `0.2 deg (%)` = round(100 * tau_key / cell_ha["0.2"], 2),
  `0.4 deg (%)` = round(100 * tau_key / cell_ha["0.4"], 2),
  `1 deg (%)`   = round(100 * tau_key / cell_ha["1"],   2))

knitr::kable(tau_tab, caption = "Detectability threshold tau as a percentage of the grid-cell area at the equator. The analysis centres on tau = 1% of the cell (about 500, 2000 and 12400 ha); extreme disagreement denotes tau_max above ten times this value.")
```

## R1.5: temporal heterogeneity does not explain the disagreement

We test whether the disagreement reflects the different moments the maps
represent. Existence of irrigation is the metric throughout (see Preliminary).

```{r r15_sameyear}

# Maps that share the 2005 reference year versus the full ensemble --------------

sameyear_2005 <- c("gmia", "gripc", "nagaraj", "meier")

data.table(
  set  = c("Same year (2005)", "Full ensemble"),
  maps = c(length(sameyear_2005), length(maps)),
  existential_pct = round(100 * c(exist_union(wide, sameyear_2005),
                                  exist_union(wide, maps))))
```

```{r r15_sizecontrol}

# Recompute over every four-map combination to hold ensemble size fixed --------

combs4  <- combn(maps, 4, simplify = FALSE)
ev4     <- sapply(combs4, function(cc) exist_union(wide, cc))
span4   <- sapply(combs4, function(cc) diff(range(nominal[cc])))
same_ev <- exist_union(wide, sameyear_2005)

data.table(
  quantity = c("Minimum", "Median", "Maximum",
               "Same-year (2005) combination", "Most dispersed (1999-2015)"),
  existential_pct = round(100 * c(min(ev4), median(ev4), max(ev4),
                                  same_ev, mean(ev4[span4 == max(span4)]))),
  percentile = c(NA, NA, NA, round(100 * mean(ev4 <= same_ev)), NA))
```

```{r r15_directional}

# Direction of disagreement: under real expansion, a recent (2015) map should
# not record zero where an older (2000-2005) map detects irrigation.

old_t    <- c("giam", "mirca2000", "gmia", "gripc")   # 2000-2005
recent_t <- c("gaez_v4", "luh2", "mirca_os")          # 2015

pooled_blind <- function(new) {
  A_old   <- as.matrix(wide[, ..old_t])
  old_any <- rowSums(A_old > 0, na.rm = TRUE) > 0
  sum(old_any & wide[[new]] == 0, na.rm = TRUE) / sum(old_any)
}

data.table(
  recent_map = c("GAEZ+2015", "LUH2", "MIRCA-OS"),
  sees_zero_where_older_detects_pct = round(100 * sapply(recent_t, pooled_blind)))
```

\newpage

# Reviewer 2

## R2.1: census products agree because they share source data, not reality

The five FAO-statistics products share definitions and much of their source
data. We compare their existential disagreement against all five-map subsets.

```{r r21}

census_ev <- exist_union(wide, census)

combs5 <- combn(maps, 5, simplify = FALSE)
ev5    <- sapply(combs5, function(cc) exist_union(wide, cc))

data.table(
  quantity = c("Census-5 (GMIA, MIRCA2000, MIRCA-OS, SPAM, GAEZ+2015)",
               "All 252 five-map subsets: minimum",
               "All 252 five-map subsets: median",
               "All 252 five-map subsets: maximum"),
  existential_pct = round(100 * c(census_ev, min(ev5), median(ev5), max(ev5))),
  note = c(sprintf("percentile %.0f%%", 100 * mean(ev5 <= census_ev)),
           "", "", ""))
```

## R2.2: remote-sensing and census products are not a subset of one another

If remote sensing simply mapped a superset of census irrigation, it would never
record zero where a census product detects irrigation. We test both directions
and the disagreement within each family.

```{r r22}

any_pos <- function(cols) rowSums(as.matrix(wide[, ..cols]) > 0, na.rm = TRUE) > 0
cP <- any_pos(census)
rP <- any_pos(rs)

data.table(
  quantity = c("RS records zero where census detects irrigation (% of census-irrigated cells)",
               "Census records zero where RS detects irrigation (% of RS-irrigated cells)",
               "Within RS family: GIAM vs GRIPC existential disagreement (%)",
               "Within census family: five FAO products existential disagreement (%)",
               "Mean global total, census maps (Mha)",
               "Mean global total, RS maps (Mha)"),
  value = c(round(100 * sum(cP & !rP) / sum(cP)),
            round(100 * sum(rP & !cP) / sum(rP)),
            round(100 * exist_union(wide, rs)),
            round(100 * exist_union(wide, census)),
            round(mean(sapply(census, function(m) sum(wide[[m]])))),
            round(mean(sapply(rs,     function(m) sum(wide[[m]]))))))
```

\newpage

# Reviewer 3

## R3.14: disagreement does not track irrigation practice

The reviewer asks whether irrigation technique (e.g. drip in Spain) drives the
disagreement. We report existential and extreme disagreement per country.

```{r r314}

# Global reference -------------------------------------------------------------

glob <- data.table(country = "GLOBAL", cells = nrow(wide),
                   existential_pct = round(100 * exist_union(wide, maps)),
                   extreme_pct = round(100 * extreme_frac(wide, maps)))

# By country -------------------------------------------------------------------

ctys <- c("India", "Portugal", "Italy", "Pakistan", "China", "Spain",
          "Egypt", "France", "Mexico", "Iran", "United States")

res <- rbindlist(lapply(ctys, function(cn) {
  s <- wide[country == cn]
  if (nrow(s) == 0) return(NULL)
  data.table(country = cn, cells = nrow(s),
             existential_pct = round(100 * exist_union(s, maps)),
             extreme_pct = round(100 * extreme_frac(s, maps)))
}))

res <- rbind(glob, res[order(-extreme_pct)])

knitr::kable(res, caption = "Existential and extreme disagreement by country (0.2 deg). Existential: of the cells where at least one map detects irrigation, the percentage in which at least one map detects none. Extreme: percentage of cropland cells whose tau_max exceeds ten times the detectability threshold. Spain (the highest-drip country) is not a maximum.")
```

\newpage

# Session information

```{r session_information}

sessionInfo()
```
